DNNNNR0: How to Avoid OOTA Without Really Trying

### 2025-03-07 paper/ootavacuous.tex
ootavacuous: Add a section 1.3 ("Our Approach")
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

All this work focused on either identifying OOTA or helping C++
implementations to avoid it.
None applied real-world hardware ordering constraints (based on
long-standing laws of physics) to the problem of avoiding OOTA cycles.

Augmenting C++ abstractions with these constraints would of course
limit applicability of this paper's results to C++ programs running
on real-world computer systems.
However, a quarter century of OOTA research based on pure C++ abstractions
has resulted in incremental progress at best, and real-world computer
systems comprise an exceedingly important part of the overall OOTA
problem.
In the wise words of Peter J.~Denning, ``Maybe we need to occasionally
descend from the high clouds of our abstractions to the concrete earthy
concerns of everyday life.''~\cite{PeterJDenning2025Abstraction}.

Therefore, this paper's contribution is to show how these
constraints prevent real-world C++ implementations doing per-thread
analysis from exhibiting OOTA behavior, and not with minimal
cost~\cite{PiezhaoOu2018OOTApapl}, but rather with no cost at
all~\cite{PaulEMcKenney2023P3046R2/OOTAwithoutTrying}.
\emph{@@@ Are these the right qualifiers?}

>>>>

All this work focused on either identifying OOTA or helping C++
implementations to avoid it.

\subsection{Our Approach}
\label{sec:Our Approach}

Our approach applies real-world hardware constraints (which are further
constrained by long-standing laws of physics) to the problem of avoiding
OOTA cycles.
We do this by augmenting the C++ abstract machine with these constraints,
which of course limits the applicability of our results to C++ programs
compiled and run on real-world computer systems.
Given that a quarter century of OOTA research based on pure unaugmented
C++ abstractions has resulted in (at best) incremental progress on the
one hand, and that real-world computer systems comprise an exceedingly
important part of the overall OOTA problem on the other, our results
are valuable despite these limitations.
In the wise words of Peter J.~Denning, ``Maybe we need to occasionally
descend from the high clouds of our abstractions to the concrete earthy
concerns of everyday life''~\cite{PeterJDenning2025Abstraction}.

The key observations that enable us to apply these constraints are:
(1)~Communicating data within a computer system takes time,
(2)~Executing instructions that compute an output value that semantically
depends on given input values also takes time,
and
(3)~Each link in an OOTA cycle consists of either data communication
or computation of an output that semantically depends on an input.
Because closing an OOTA cycle comprising a sequence of commmunications
and computations requires this sequence to end at the same time as it
starts, OOTA cycles cannot form in those C++ implementations that are
within the scope of this paper.
A great advantage of our approach is that we are \emph{not} required to
precisely define the boundaries of sdep.
After all, although compilers are forbidden from breaking semantic
dependencies, they are by no means required to break non-semantic
dependencies.

These simple-seeming observations conceal considerable complexity,
and so much of the remainder of this paper deals with corner cases that
arise when carefully considering semantic dependencies, executions, C++
compilers, and hardware properties.

In short, this paper's contribution is to show how these
constraints prevent real-world C++ implementations doing per-thread
analysis from exhibiting OOTA behavior, and not with minimal
cost~\cite{PiezhaoOu2018OOTApapl}, but rather with no cost at
all~\cite{PaulEMcKenney2023P3046R2/OOTAwithoutTrying}.
\emph{@@@ Are these the right qualifiers?}
